---
title: "Group 7 final project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(broom)
library(dplyr)
library(patchwork)
library("leaps")
library(faraway)
library(caret)
library(boot)
library(MPV) 
library(broom)
library("leaps")
library(boot)
library(ggplot2)
library("gridExtra")
library(mgcv)
library(gamclass)
```

## Exploratory of variables
```{r}
cancer_df = read_csv("./data/Cancer_Registry.csv") %>% 
  janitor::clean_names() %>% 
  select(target_death_rate, everything()) %>% 
  separate(geography, c("county", "state"), sep = ",") %>% 
  mutate(county = as.factor(county), 
         state = as.factor(state),
         pct_case_count = avg_ann_count / pop_est2015*100000) %>% 
  select(target_death_rate, pct_case_count, everything())
  
```

Percentage of annul case dignosed count plot
```{r}
plot_count_pct =
  cancer_df %>% 
  ggplot(aes(y = pct_case_count, x = target_death_rate, color = state)) +
  geom_point() 
  #geom_smooth(se = F)
ggplotly(plot_count_pct)
```

Incidence rate plot
```{r}
plot_incidence = 
  cancer_df %>% 
  ggplot(aes(x = incidence_rate, y = target_death_rate, color = state)) +
  geom_point() 
  #geom_smooth(se = F)
  ggplotly(plot_incidence)
  
# Influential points in the dataset, state Flordia and Virginia.
```

Income plot
```{r}
plot_income = 
  cancer_df %>% 
  ggplot(aes(x = med_income, y = target_death_rate, color = state)) +
  geom_point() 
  #geom_smooth(se = F)
  ggplotly(plot_income)
```

Age plots
```{r}
plot_age_1 =
  cancer_df %>% 
  ggplot(aes(x = median_age, y = target_death_rate)) +
  geom_point() +
  geom_smooth(se = F)
ggplotly(plot_age_1)
# error data in this column, larger than 100

cancer_df %>%
  filter(median_age < 100) %>% 
  ggplot(aes(x = median_age)) +
  geom_histogram()

plot_age_2 =
  cancer_df %>% 
  ggplot(aes(x = median_age_male, y = target_death_rate)) +
  geom_point() +
  geom_smooth(se = F)
ggplotly(plot_age_2)

plot_age_3 =
  cancer_df %>% 
  ggplot(aes(x = median_age_female, y = target_death_rate)) +
  geom_point() +
  geom_smooth(se = F)
ggplotly(plot_age_3)
```

```{r}
cancer_df %>% 
  select(-county, -state, -binned_inc) %>% 
  cor() %>% 
  as.tibble() 
```

```{r}
lm(target_death_rate ~ avg_ann_count * pop_est2015, data = cancer_df) %>% 
  summary()
```



```{r}
raw_data <- read_csv("./Data/Cancer_Registry.csv")
raw_data = janitor::clean_names(raw_data)
raw_data1 = raw_data %>% 
  select(-binned_inc) %>% 
  separate(geography, c("county", "state"), sep = ",") %>% 
  mutate(county = as.factor(county), 
         state = as.factor(state),
         avg_ann_count = avg_ann_count / pop_est2015*100000) %>% 
  select(target_death_rate, avg_ann_count, everything())

res <- cor(raw_data1 %>% select(-county, -state))
round(res, 3)
```

```{r}
summary(raw_data1$pct_employed16_over) # contains 152 NA's
summary(raw_data1$pct_unemployed16_over)
summary(raw_data1$pct_private_coverage)
summary(raw_data1$pct_private_coverage_alone) # get rid of this variable! Contains 609 NA's
summary(raw_data1$pct_emp_priv_coverage)
summary(raw_data1$pct_public_coverage)
summary(raw_data1$pct_public_coverage_alone)

histogram1 = ggplot(raw_data1, aes(x  = raw_data1$pct_employed16_over)) + geom_histogram() + labs(title = "Percent ages 16+ employed")


histogram2 = ggplot(raw_data1, aes(x  = raw_data1$pct_unemployed16_over)) + geom_histogram() + labs(title = "Percent ages 16+ unemployed")


histogram3 = ggplot(raw_data1, aes(x  = raw_data1$pct_private_coverage)) + geom_histogram() + labs(title = "Percent private health coverage")


histogram4 = ggplot(raw_data1, aes(x  = raw_data1$pct_private_coverage_alone)) + geom_histogram() + labs(title = "Percent private health coverage alone")


histogram5 = ggplot(raw_data1, aes(x  = raw_data1$pct_emp_priv_coverage)) + geom_histogram() + labs(title = "Percent employee-provided private health coverage")


histogram6 = ggplot(raw_data1, aes(x  = raw_data1$pct_public_coverage)) + geom_histogram() + labs(title = "Percent government-provided health coverage")


histogram7 = ggplot(raw_data1, aes(x  = raw_data1$pct_public_coverage_alone)) + geom_histogram() + labs(title = "Percent government-provided health coverage alone")



(histogram1 + histogram2)
(histogram3 + histogram4)
(histogram5 + histogram6)/(histogram7)

reg1 = lm(target_death_rate~pct_employed16_over, data = raw_data1)
reg2 = lm(target_death_rate~pct_unemployed16_over, data = raw_data1)
reg3 = lm(target_death_rate~pct_private_coverage, data = raw_data1)
reg4 = lm(target_death_rate~pct_private_coverage_alone, data = raw_data1)
reg5 = lm(target_death_rate~pct_emp_priv_coverage, data = raw_data1)
reg6 = lm(target_death_rate~pct_public_coverage, data = raw_data1)
reg7 = lm(target_death_rate~pct_public_coverage_alone , data = raw_data1)

plot(raw_data1$pct_employed16_over, raw_data1$target_death_rate)
abline(reg1,lwd = 2,col = 2)
plot(raw_data1$pct_unemployed16_over, raw_data1$target_death_rate)
abline(reg2,lwd = 2,col = 2)
plot(raw_data1$pct_private_coverage, raw_data1$target_death_rate)
abline(reg3,lwd = 2,col = 2)
plot(raw_data1$pct_private_coverage_alone, raw_data1$target_death_rate)
abline(reg4,lwd = 2,col = 2)
plot(raw_data1$pct_emp_priv_coverage, raw_data1$target_death_rate)
abline(reg5,lwd = 2,col = 2)
plot(raw_data1$pct_public_coverage, raw_data1$target_death_rate)
abline(reg6,lwd = 2,col = 2)
plot(raw_data1$pct_public_coverage_alone, raw_data1$target_death_rate)
abline(reg7,lwd = 2,col = 2)

```

```{r}
raw_data2 = raw_data1 %>% 
  #mutate(interaction1 = med_income * poverty_percent) %>% 
  select(target_death_rate, avg_ann_count, incidence_rate, med_income, poverty_percent, median_age_male, median_age_female, pct_bach_deg25_over, pct_unemployed16_over, pct_public_coverage_alone)

#############################################################################
# Backward elimination: take out non-significant variables 'one at a time'  #
# starting with the highest p-value                                         #
#############################################################################

reg = lm(target_death_rate~., data = raw_data2)
summary(reg)

# No Median age of female county residents
step1<-update(reg, . ~ . -median_age_female)
summary(step1)

# No Median income per county
step2<-update(step1, . ~ . -med_income)
summary(step2)

# No Mean number of reported cases of cancer diagnosed annually
step3<-update(step2, . ~ . -avg_ann_count)
summary(step3)

# No Median age of male county residents
step4<-update(step3, . ~ . -median_age_male)
summary(step4)


############################################################################
#  Stepwise regreession                                                    #   
# 'Step' function uses AIC criterion for var selection and the default     #
#  option is 'backward'.                                                   #
#  Step is a simplified version of stepAIC()                               #
############################################################################


mult.fit <- lm(target_death_rate~., data = raw_data2)
step(mult.fit, direction = 'both')


#########################################################################################
# Forward elimination: Reversed backward elimination starting with the lowest p-value   #                                                            
#########################################################################################


### Step 1:  Fit simple linear regressions for all variables,look for the variable with lowest p-value
fit1 <- lm(target_death_rate ~ avg_ann_count, data=raw_data2)
tidy(fit1)
fit2 <- lm(target_death_rate ~ incidence_rate, data=raw_data2)
tidy(fit2)
fit3 <- lm(target_death_rate ~ med_income, data=raw_data2)
tidy(fit3)
fit4 <- lm(target_death_rate ~ poverty_percent, data=raw_data2)
tidy(fit4)
fit5 <- lm(target_death_rate ~ median_age_male, data=raw_data2)
tidy(fit5)
fit6 <- lm(target_death_rate ~ median_age_female, data=raw_data2)
tidy(fit6)
fit7 <- lm(target_death_rate ~ pct_bach_deg25_over, data=raw_data2)
tidy(fit7)
fit8 <- lm(target_death_rate ~ pct_unemployed16_over, data=raw_data2)
tidy(fit8)
fit9 <- lm(target_death_rate ~ pct_public_coverage_alone, data=raw_data2)
tidy(fit9)


# Enter first the one with the lowest p-value: pct_bach_deg25_over
forward1<-lm(target_death_rate~pct_bach_deg25_over , data=raw_data2)
tidy(forward1)

### Step 2: Enter the one with the lowest p-value in the rest 
fit1 <- update(forward1, . ~ . +avg_ann_count)
tidy(fit1)
fit2 <- update(forward1, . ~ . +incidence_rate)
tidy(fit2)
fit3 <- update(forward1, . ~ . +med_income)
tidy(fit3)
fit4 <- update(forward1, . ~ . +poverty_percent)
tidy(fit4)
fit5 <- update(forward1, . ~ . +median_age_male)
tidy(fit5)
fit6 <- update(forward1, . ~ . +median_age_female)
tidy(fit6)
fit7 <- update(forward1, . ~ . +pct_unemployed16_over)
tidy(fit7)
fit8 <- update(forward1, . ~ . +pct_public_coverage_alone)
tidy(fit8)



# Enter the one with the lowest p-value: incidence_rate
forward2 <- update(forward1, . ~ . + incidence_rate)
tidy(forward2)

### Step 3: Enter the one with the lowest p-value in the rest 
fit1 <- update(forward2, . ~ . +avg_ann_count)
tidy(fit1)
fit2 <- update(forward2, . ~ . +med_income)
tidy(fit2)
fit3 <- update(forward2, . ~ . +poverty_percent)
tidy(fit3)
fit4 <- update(forward2, . ~ . +median_age_male)
tidy(fit4)
fit5 <- update(forward2, . ~ . +median_age_female)
tidy(fit5)
fit6 <- update(forward2, . ~ . +pct_unemployed16_over)
tidy(fit6)
fit7 <- update(forward2, . ~ . +pct_public_coverage_alone)
tidy(fit7)


# Enter the one with the lowest p-value: poverty_percent
forward3 <- update(forward2, . ~ . + poverty_percent)
tidy(forward3)

### Step 4: Enter the one with the lowest p-value in the rest 
fit1 <- update(forward3, . ~ . +avg_ann_count)
tidy(fit1)
fit2 <- update(forward3, . ~ . +med_income)
tidy(fit2)
fit3 <- update(forward3, . ~ . +median_age_male)
tidy(fit3)
fit4 <- update(forward3, . ~ . +median_age_female)
tidy(fit4)
fit5 <- update(forward3, . ~ . +pct_unemployed16_over)
tidy(fit5)
fit6 <- update(forward3, . ~ . +pct_public_coverage_alone)
tidy(fit6)


# Enter the one with the lowest p-value: pct_unemployed16_over 
forward4 <- update(forward3, . ~ . + pct_unemployed16_over )
tidy(forward4)


### Step 5: Enter the one with the lowest p-value in the rest 
fit1 <- update(forward4, . ~ . +avg_ann_count)
tidy(fit1)
fit2 <- update(forward4, . ~ . +med_income)
tidy(fit2)
fit3 <- update(forward4, . ~ . +median_age_male)
tidy(fit3)
fit4 <- update(forward4, . ~ . +median_age_female)
tidy(fit4)
fit5 <- update(forward4, . ~ . +pct_public_coverage_alone)
tidy(fit5)

# Enter the one with the lowest p-value: pct_public_coverage_alone
forward5 <- update(forward4, . ~ . + pct_public_coverage_alone )
tidy(forward5)


### Step 6: Enter the one with the lowest p-value in the rest 
fit1 <- update(forward5, . ~ . +avg_ann_count)
tidy(fit1)
fit2 <- update(forward5, . ~ . +med_income)
tidy(fit2)
fit3 <- update(forward5, . ~ . +median_age_male)
tidy(fit3)
fit4 <- update(forward5, . ~ . +median_age_female)
tidy(fit4)


# P-value of all new added variables are larger than 0.15, which means that they 
# are not significant predictor, and we stop here.

# The model we obtained is target_death_rate~population + murder + hs_grad + frost
mult.fit <- lm(target_death_rate~pct_bach_deg25_over + incidence_rate + poverty_percent + pct_public_coverage_alone + pct_unemployed16_over , data = raw_data2)
summary(mult.fit)
```

## Predictive capability of the model

###K-fold Cross Validation (CV)
##### Use 5-fold validation and create the training sets
```{r}
set.seed(2)
# Use 5-fold validation and create the training sets

data_train <- trainControl(method = "cv", number = 5)

# Fit the 4-variables model
model_caret <- train(target_death_rate~pct_bach_deg25_over + incidence_rate + poverty_percent + pct_public_coverage_alone + pct_unemployed16_over, data = raw_data2,
                   trControl = data_train,
                   method = 'lm',
                   na.action = na.pass)
  
# Model predictions using 9 parts of the data fro training 
model_caret

#Rsquared - good % variantion acconted for

# Model coefficients
model_caret$finalModel

# Examine model prediction for each fold
model_caret$resample

# Look at standard deviation around the Mean Square Error value by examining the Root Mean Square Error from each fold.
sd(model_caret$resample$RMSE^2)

```

##### Use 10-fold validation and create the training sets
```{r}
set.seed(2)
# Use 10-fold validation and create the training sets

data_train1 <- trainControl(method = "cv", number = 10)

# Fit the 4-variables model
model_caret1 <- train(target_death_rate~pct_bach_deg25_over + incidence_rate + poverty_percent + pct_public_coverage_alone + pct_unemployed16_over, data = raw_data2,
                   trControl = data_train1,
                   method = 'lm',
                   na.action = na.pass)
  
# Model predictions using 9 parts of the data fro training 
model_caret1

#Rsquared - good % variantion acconted for

# Model coefficients
model_caret1$finalModel

# Examine model prediction for each fold
model_caret1$resample

# Look at standard deviation around the Mean Square Error value by examining the Root Mean Square Error from each fold.
sd(model_caret1$resample$RMSE^2)



# Let's look at the model using all data, no CV
full_model <- lm(target_death_rate~pct_bach_deg25_over + incidence_rate + poverty_percent + pct_public_coverage_alone + pct_unemployed16_over , data = raw_data2)
summary(full_model)

# Full data had an Rsquared=0.4704
```

```{r}
summary(model_caret$resample$RMSE^2)
summary(model_caret1$resample$RMSE^2)


par(mfrow = c(1,2))
boxplot(model_caret$resample$RMSE^2, main = "MSE of 5 fold CV", ylim = c(330, 500))
boxplot(model_caret1$resample$RMSE^2, main = "MSE of 10 fold CV", ylim = c(330, 500))
```


### Bootstrap to assess the variability of model estimates: b0 and b1
```{r}
boot.fn<-function(data, index){
	return(coef(lm(target_death_rate~pct_bach_deg25_over + incidence_rate + poverty_percent + pct_public_coverage_alone + pct_unemployed16_over , data = raw_data2, subset=index)))}

# Our usual regression, no bootstrap yet
boot.fn(raw_data2,1:3047)


# Compute the estimates by sampling with replacement
# Sample chooses 3047 observations from 3047, with replacement
# Might have duplicates
set.seed(1)

# One draw
boot.fn(raw_data2,sample(3047,3047,replace=T))

# Use function boot() to repeat the sampling 10000 times.
# Repeat 10000 times to get the estimates, SEs ad bias

boot(raw_data2, boot.fn, 10000)

```


###Model validation: criteria  
```{r}
newsummary <- function(model)
{
    list('coefs'    = round(t(summary(model)$coef[, 1:2]), 4),
         'criteria' = cbind('SSE'   = anova(model)["Residuals", "Sum Sq"],
                            'PRESS' = PRESS(model),
                            'MSE'   = anova(model)["Residuals", "Mean Sq"],
                            'Rsq'   = summary(model)$adj.r.squared))
}

newsummary(lm(target_death_rate~pct_bach_deg25_over + incidence_rate + poverty_percent + pct_public_coverage_alone + pct_unemployed16_over , data = raw_data2))
```

```{r}
raw_data <- read_csv("./Data/Cancer_Registry.csv")
raw_data = janitor::clean_names(raw_data)
raw_data1 = raw_data %>% 
  select(-binned_inc) %>% 
  separate(geography, c("county", "state"), sep = ",") %>% 
  mutate(county = as.factor(county), 
         state = as.factor(state),
         avg_ann_count = avg_ann_count / pop_est2015*100000) %>% 
  select(target_death_rate, avg_ann_count, everything())
```

```{r}
sapply(raw_data1, function(x) sum(is.na(x)))
```

```{r}
## Choose possible variables for our model
raw_data2 = raw_data1 %>% 
  select(target_death_rate, incidence_rate, poverty_percent,  pct_bach_deg25_over, pct_unemployed16_over, pct_public_coverage_alone)
sapply(raw_data2, function(x) sum(is.na(x)))
## No missing values
```

```{r}
## Visualization of predictors to find outliers
histogram1 = ggplot(raw_data2, aes(x  = raw_data2$target_death_rate)) + geom_histogram() + labs(title = "pct_target_death_rate")
histogram2 = ggplot(raw_data2, aes(x  = raw_data2$incidence_rate)) + geom_histogram() + labs(title = "pct_incidence_rate")
histogram3 = ggplot(raw_data2, aes(x  = raw_data2$poverty_percent)) + geom_histogram() + labs(title = "pct_poverty_percent")
histogram4 = ggplot(raw_data2, aes(x  = raw_data2$pct_bach_deg25_over)) + geom_histogram() + labs(title = "pct_pct_bach_deg25_over")
histogram5 = ggplot(raw_data2, aes(x  = raw_data2$pct_unemployed16_over)) + geom_histogram() + labs(title = "pct_unemployed16_over")
histogram6 = ggplot(raw_data2, aes(x  = raw_data2$pct_public_coverage_alone)) + geom_histogram() + labs(title = "pct_public_coverage_alone")
grid.arrange(histogram1, histogram2, histogram3, histogram4, histogram5,histogram6, ncol=2)
histogram1 = ggplot(raw_data2, aes(x  = raw_data2$target_death_rate)) + geom_point() + labs(title = "pct_target_death_rate")
```

```{r}
## dotplot to make outliers more distinct
dot1 = ggplot(raw_data2, aes(x  = raw_data2$target_death_rate, y =raw_data2$target_death_rate)) + geom_point() + labs(title = "pct_target_death_rate")
dot2 = ggplot(raw_data2, aes(x  = raw_data2$incidence_rate, y =raw_data2$target_death_rate)) + geom_point() + labs(title = "pct_incidence_rate")
dot3 = ggplot(raw_data2, aes(x  = raw_data2$poverty_percent, y =raw_data2$target_death_rate)) + geom_point() + labs(title = "pct_poverty_percent")
dot4 = ggplot(raw_data2, aes(x  = raw_data2$pct_bach_deg25_over, y =raw_data2$target_death_rate)) + geom_point() + labs(title = "pct_pct_bach_deg25_over")
dot5 = ggplot(raw_data2, aes(x  = raw_data2$pct_unemployed16_over, y =raw_data2$target_death_rate)) + geom_point() + labs(title = "pct_unemployed16_over")
dot6 = ggplot(raw_data2, aes(x  = raw_data2$pct_public_coverage_alone, y =raw_data2$target_death_rate)) + geom_point() + labs(title = "pct_public_coverage_alone")
grid.arrange(dot1, dot2, dot3, dot4, dot5, dot6, ncol=2)
```



```{r}
## take out outliers 
raw_data3 = raw_data2[-c(1490),]
raw_data3 = raw_data3[-c(282),]
## Two rows are outliers
```

```{r}
## Plot new datasets with no outliers
dot1 = ggplot(raw_data3, aes(x  = raw_data3$target_death_rate, y =raw_data3$target_death_rate)) + geom_point() + labs(title = "pct_target_death_rate")
dot2 = ggplot(raw_data3, aes(x  = raw_data3$incidence_rate, y =raw_data3$target_death_rate)) + geom_point() + labs(title = "pct_incidence_rate")
dot3 = ggplot(raw_data3, aes(x  = raw_data3$poverty_percent, y =raw_data3$target_death_rate)) + geom_point() + labs(title = "pct_poverty_percent")
dot4 = ggplot(raw_data3, aes(x  = raw_data3$pct_bach_deg25_over, y =raw_data3$target_death_rate)) + geom_point() + labs(title = "pct_pct_bach_deg25_over")
dot5 = ggplot(raw_data3, aes(x  = raw_data3$pct_unemployed16_over, y =raw_data3$target_death_rate)) + geom_point() + labs(title = "pct_unemployed16_over")
dot6 = ggplot(raw_data3, aes(x  = raw_data3$pct_public_coverage_alone, y =raw_data3$target_death_rate)) + geom_point() + labs(title = "pct_public_coverage_alone")
grid.arrange(dot1, dot2, dot3, dot4, dot5, dot6, ncol=2)
```

```{r}
## Descriptive statistics
summary(raw_data3)
```

```{r}
mult.fit1 <- lm(target_death_rate~pct_bach_deg25_over + incidence_rate + poverty_percent + pct_public_coverage_alone + pct_unemployed16_over , data = raw_data2)
summary(mult.fit1)
## Cook distance
stu_res<-rstandard(mult.fit1)
outliers_y<-stu_res[abs(stu_res)>2.5]

# Measures of influence:
# Gives DFFITS, Cook's Distance, Hat diagonal elements, and others.

influence.measures(mult.fit1)

# Look at the Cook's distance lines and notice obs 5 and 28 as potential Y outliesrs / influential points

par(mfrow=c(2,2))
plot(mult.fit1)
```

## We have potential outliers, however the outlier is not influential so we decide to keep it.


```{r}
gam1 <- gam(target_death_rate~s(pct_bach_deg25_over) + s(incidence_rate) + s(poverty_percent) + s(pct_public_coverage_alone) + s(pct_unemployed16_over) , data = raw_data2)
gam2 <- gam(target_death_rate~s(pct_bach_deg25_over) + s(incidence_rate) + s(poverty_percent) + s(pct_public_coverage_alone) + s(pct_unemployed16_over) , family=Gamma(link=log), data = raw_data2)
summary(gam1)
summary(gam2)
par(mfrow = c(1,5))
plot(gam1)
gam3 <- gam(target_death_rate~s(pct_bach_deg25_over) + s(incidence_rate) + s(poverty_percent) + s(pct_public_coverage_alone) + s(pct_unemployed16_over) , method = "GACV.Cp",data = raw_data2)
summary(gam3)
gam4 <- gam(target_death_rate~s(pct_bach_deg25_over) + s(incidence_rate) + s(poverty_percent) + s(pct_public_coverage_alone) + s(pct_unemployed16_over) + s(med_income) +s(pct_white) +s(pct_white)+s(pct_black)+s(pct_asian) +s(median_age_female), data = raw_data1)
summary(gam4)
```

```{r}
RSS <- c(crossprod(gam2$residuals))
MSE <- RSS / length(gam2$residuals)
RSS1 <- c(crossprod(gam4$residuals))
MSE1 <- RSS / length(gam4$residuals)
MSE1

ggplot(raw_data3, aes(x = poverty_percent, y = target_death_rate)) + 
  geom_point() + 
  labs(title = "pct_poverty_percent", 
       x = "Incident rate", 
       y = "Target death")
```


